---
title: "Review of Air Quality App Usage"
runningheader: "Air Quality App Usage" # only for pdf output
author: "David Hall"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_html: default
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
library(lubridate)
library(tidyverse)
library(cowplot)
library(scales)
# invalidate cache when the tufte version changes
#knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
#options(htmltools.dir.version = FALSE)
```

# Background

The CHM135 Winter 2020 Practical 1 has wrapped up, and I downgraded our server status to the base (free) option. Given that it'll be a while until we get student feedback, I though I would comb through the sparse shiny apps metrics and see if I could get a better idea of how the app was used. 

# Results and Discussion

There isn't much recorded by the Shiny server in terms of user metrics; the metrics that are recorded relate to server usage and not user experience. A useful metric is "connections", as in "when a person goes to the app website". Note that there can be multiple connections for a given individual (i.e. they go to the website on two different days). Nonetheless, it provides a useful approximate for the number of people who've used the app. Figure 1 shows the app connections over time. Predictably, the most connection are towards the tail end of the *Practical 1* synchronous sessions (Jan 18th to 22rd), and peaking over the weekend. 

```{r app-connections-data, eval = FALSE, echo = FALSE, warning = FALSE, error = FALSE, message=FALSE}

#Code modified from Simon Woodward, Stack Overflow; https://stackoverflow.com/a/53733859

df <- rsconnect::showMetrics("container_status",
                             c("connect_count",
                               "connect_procs"),
                             appName="AirQualityApp",
                             server="shinyapps.io",
                             from="14w",
                             interval="1m"
)

colnames(df) <- c("connect_count", "metricSeries", "connect_procs", "timestamp" )

#write.csv(df, file = "container_status_Practical1.csv")

```



```{r app-connections_data-manipulation, echo = FALSE, error = FALSE, message = FALSE, warning = FALSE}

# Container data spanning lab period

df <- read.csv("./dataForPaper/container_status_Practical1.csv") %>% 
  mutate(date=as.POSIXct(as.numeric(as.character(timestamp)),origin="1970-01-01")) %>% 
  mutate(connect_count = as.numeric(connect_count)) %>%
  mutate(connect_procs = as.numeric(connect_procs))

# Subsetting data to cover Exp 1 period & getting number of connections

df1 <- df %>%
  select(-timestamp) %>% 
  filter(date > ymd_hms("2021-01-10 00:00:00") & date < ymd_hms("2021-01-31 00:00:00")) %>%
  arrange(date) %>% 
  mutate(
    n_count=cumsum(connect_count),
    n_procs=cumsum(connect_procs),
    new_connect=case_when(
      connect_count>lag(connect_count,1) ~ connect_count-lag(connect_count,1),
      TRUE ~ 0),
    n_connect=cumsum(new_connect) # approximate
  ) %>% 
  filter(n_count>0)

# Prep for cumsum plot

df2 <- df1 %>%  
  select(n_connect, date) %>% 
  gather(key="key", value="value", -date)

# getting counts per day for bar plot.

df3 <- df1 %>%
  mutate(date2 = as.Date(date)) %>%
  group_by(date2) %>%
  summarize(size = max(n_connect)) %>%
  mutate(dayCount = size - lag(size))

```

```{r app-connections, fig.height = 2.5, echo = FALSE, error = FALSE, message = FALSE, warning = FALSE, fig.cap="Connections to Air Quality app per day (top) and cummulative connections over time (bottom). Vertical lines indicate when the lab material was made available online to students, when the synchronous sessions with TAs started, and the due date. Blue rectangle highlights period when students worked on report sheet."}

p1 <- ggplot(df3, aes(x = date2, y = dayCount)) +
  geom_bar(fill="gray", stat = "identity") +  theme(axis.title=element_blank()) +
  scale_y_continuous(breaks=seq(0, 70, 20)) +
  labs(y = "Connections\nper day") +
  theme_classic() +
  theme(axis.line.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x=element_blank(),
        panel.grid.minor.x=element_blank(),
        panel.grid.major.x=element_blank())
  

p2 <- ggplot(df2) +
  labs(x="Date", y="Cumulative\nConnections") +
  geom_line(aes(x=date, y=value)) + 
  geom_vline(xintercept = ymd_hms("2021-01-10 18:00:00"), color = "black", lwd = 1) + #lab published on quercus
  geom_vline(xintercept = ymd_hms("2021-01-21 09:00:00"), color = "black", lwd = 1) + #Sync sessions starts
  geom_vline(xintercept = ymd_hms("2021-01-27 22:00:00"), color = "black", lwd = 1) + #due date 
  annotate("rect", xmin = ymd_hms("2021-01-21 09:00:00"), xmax = ymd_hms("2021-01-27 22:00:00"), 
           ymin = 0, ymax = max(df2$value), alpha = .1,fill = "blue") + # rect for when students actively working
  annotate("text", x = ymd_hms("2021-01-11 00:00:00"), y = 250, label = "Released", hjust = 0, size=3) +
  annotate("text", x = ymd_hms("2021-01-21 00:00:00"), y = 125, label = "Start sync. \nsessions", hjust = 1, size=3) +
  annotate("text", x = ymd_hms("2021-01-27 23:59:00"), y = 125, label = "Due date", hjust = 0, size=3) +
  scale_x_datetime(date_breaks = "7 day", date_labels = "%b %d") +
  theme_classic()




#grid.arrange(p2, p, ncol = 1)
plot_grid(p1, p2, ncol = 1, align = "v")

```

## Server time

From the Shiny website, **we've used 77.8 hrs** over the course of the lab period. With **260 connections**, it means that **each connection (read user session) lasted approximately 18 minutes**. Some of these connection are ours (i.e. me checking on the app), but we also didn't spend too long looking at elements on the app, so I believe the 18 minutes/student is a reasonable measurement of student engagement with the App. Note, this is the time that a person was on the website, and not the amount of time the servers spent running calculations and fetching data. 

| Metric            | Numbers  |
|-------------------|----------|
| Total Connections | 260      |
| Server time       | 77.8 hrs |
| Time/connection   | 18 mins  |

Table: Summary of student engagement (approximate)

## Fall 2021 Metrics

The lab was run over the fall term with over 2000 students enrolled in the course. 

```{r app-connections-data, eval = FALSE, echo = FALSE, warning = FALSE, error = FALSE, message=FALSE}

#Code modified from Simon Woodward, Stack Overflow; https://stackoverflow.com/a/53733859

df <- rsconnect::showMetrics("container_status",
                             c("connect_count",
                               "connect_procs"),
                             appName="AirQualityApp",
                             server="shinyapps.io",
                             from="6w",
                             interval="1m"
)

colnames(df) <- c("connect_count", "metricSeries", "connect_procs", "timestamp" )

#write.csv(df, file = "./dataForPaper/Fall2021.csv")

```



```{r app-connections_data-manipulation, echo = FALSE, error = FALSE, message = FALSE, warning = FALSE}

# Container data spanning lab period

fall <- read.csv("./dataForPaper/Fall2021.csv") %>% 
  mutate(date=as.POSIXct(as.numeric(as.character(timestamp)),origin="1970-01-01")) %>% 
  mutate(connect_count = as.numeric(connect_count)) %>%
  mutate(connect_procs = as.numeric(connect_procs))

# Subsetting data to cover Exp 1 period & getting number of connections

fall1 <- fall %>%
  select(-timestamp) %>% 
  filter(date > ymd_hms("2021-09-17 00:00:00") & date < ymd_hms("2021-10-15 00:00:00")) %>%
  arrange(date) %>% 
  mutate(
    n_count=cumsum(connect_count),
    n_procs=cumsum(connect_procs),
    new_connect=case_when(
      connect_count>lag(connect_count,1) ~ connect_count-lag(connect_count,1),
      TRUE ~ 0),
    n_connect=cumsum(new_connect) # approximate
  ) %>% 
  filter(n_count>0)

# Prep for cumsum plot

fall2 <- fall1 %>%  
  select(n_connect, date) %>% 
  gather(key="key", value="value", -date)

# getting counts per day for bar plot.

fall3 <- fall1 %>%
  mutate(date2 = as.Date(date)) %>%
  group_by(date2) %>%
  summarize(size = max(n_connect)) %>%
  mutate(dayCount = size - lag(size))

```

```{r app-connections, fig.height = 2.5, echo = FALSE, error = FALSE, message = FALSE, warning = FALSE, fig.cap="Connections to Air Quality app per day (top) and cummulative connections over time (bottom). Vertical lines indicate when the lab material was made available online to students, when the synchronous sessions with TAs started, and the due date. Blue rectangle highlights period when students worked on report sheet."}

p1 <- ggplot(fall3, aes(x = date2, y = dayCount)) +
  geom_bar(fill="gray", stat = "identity") +  theme(axis.title=element_blank()) +
  labs(y = "Connections\nper day") +
  theme_classic() +
  theme(axis.line.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x=element_blank(),
        panel.grid.minor.x=element_blank(),
        panel.grid.major.x=element_blank())
  

p2 <- ggplot(fall2) +
  labs(x="Date", y="Cumulative\nConnections") +
  geom_line(aes(x=date, y=value)) + 
  scale_x_datetime(date_breaks = "7 day", date_labels = "%a \n%b %d") +
  theme_classic()




#grid.arrange(p2, p, ncol = 1)
plot_grid(p1, p2, ncol = 1, align = "v")

```

Over 3000 students signed on, probably going back for more.
