---
title: "Tidying ECCC Air Quality Data"
author: "David Hall"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_html: default
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

# Source

Air quality data is taken from Environment and Climate Change Canada's (ECCC) National Air Pollution Surveillance Program (NAPS). This specific applet uses *annual continuous hourly measurements*. Each of these files contains the hourly measurements of a single pollutant (i.e. O~3~) across all operational NAPS station for the given year.[^1](<http://data.ec.gc.ca/data/air/monitor/national-air-pollution-surveillance-naps-program/Data-Donnees/2018/ContinuousData-DonneesContinu/?lang=en>)] There is at least a year delay between collection and publication of data.[^2]

[^1]: Files can be found on the ECCC website here

[^2]: At the time of writing, the most recent annual report is from 2018

## Data organization

Each yearly pollutant dataset can be downloaded as a `.csv` from the ECCC website. As listed on the dataset preamble:

-   Measurements are reported in parts-per-billion (ppb)
-   Data is ending local standard time. (i.e. H01 is the hourly measurements from 00:00 to 01:00).
-   Zeros are valid values
-   `-999` denotes no data available.

However, the structure of this dataset have some incompatible elements with the `tidyverse` ecosystem, these include:

-   Matrix style layout, with each row corresponding to a day, and columns for each hourly report.
-   Bilingual headers separated by "//"
-   Separation of *date* and *time*

## Why i'm not storing server data as tidy

```{=tex}
\begin{marginfigure}
  \includegraphics[width=\linewidth]{./images/widevtidy.png} 
  \caption{Wide vs. tidy data layout in flat files (i.e. csv).}
  \label{widevtidy}
\end{marginfigure}
```
My first idea was to combine two ECCC annual datasets into a single file, tidy, file. While R has some tricks to handling these types of files in memory, when it comes to actually saving, the resulting file was exceptionally large. For example, combining the 10.3 mb `O3_2018.csv` with the 7.5 mb `NO2_2018.csv` resulted in a 126 mb `.csv` file. Correspondingly, I'm going to look for an alternative way to store and access the data. 

I'm currently thinking about combining datasets in their original 'wide' format, and reactively subsetting and tidying them on the fly as users input calculations. 

# Converting ECCC NAPS data for Applet

Packages used include:

```{r, message = FALSE}
library(tidyverse)
library(anytime) # Quicker than tiverse's lubridate package
```

# Importing and joining ECCC datasets

The goal is to combine multiple ECCC hourly datasets in their 'wide' format. In other words, so that each row is a day, and the hourly data is stored as a column.  

```{r, eval=FALSE, error=FALSE, warning=FALSE, message=FALSE}

# Adding pollutant as prefix to hour columns. 
ECCCcolRename <- function(df, chem){
   
  df <- df %>% rename_all(funs(gsub("\\..*","", make.names(names(df))))) %>%
    rename_at(vars(starts_with("H")), ~paste0(chem,"_", .))
  
}


# Only worrying about O3 and NO2 for now
combineO3andNO2 <- function(O3, NO2, nrows = Inf){
  
  O3dat <-  read.csv(O3, skip = 7, nrows = nrows) %>%
    ECCCcolRename(., "O3") %>%
    select(-c("Pollutant"))
    
  NO2dat <- read.csv(NO2, skip = 7, nrows = nrows) %>%
    ECCCcolRename(., "NO2") %>%
    select(-c("Latitude", "Longitude", "P", "City", "Pollutant"))
  
  df <- O3dat %>%
    inner_join(NO2dat, by = c("NAPS", "Date")) %>%
    na_if(-999)
    
}

datWide <- combineO3andNO2(O3 = "O3_2018.csv", NO2 = "NO2_2018.csv")

#write.csv(datWide, file = "www/ECCC2018_wideCombined.csv")

```

## Classifying population centres

A tentative question is for students to compare their urban (downtown Toronto) data to that of a rural/smaller population center. To this end, i'll need to identify the population of each NAPS ID assosciated City and use that vale to colour the map marker. StatsCan *Population and Dwelling Count Highlight Tables, 2016 Census* has population counts from all census populations. Data can be found [here](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/hlt-fst/pd-pl/comprehensive.cfm) According to Stats Canada an "'urban area' was defined as having a population of at least 1,000 and a density of 400 or more people per square kilometre. All territory outside an urban area was defined as rural area."

They also offer classification on urban population centres:

- **small population centres**, with a population of between 1,000 and 29,999
- **medium population centres**, with a population of between 30,000 and 99,999
- **large urban population centres**, consisting of a population of 100,000 and over

```{r}
pops <- read.csv("NAPSPops.csv")

# Adding population to wide data
df <- datWide %>% 
inner_join(pops, by = "City") %>%
  mutate(PopSize = if_else(Population >= 100000, "large",
                           if_else(Population < 100000 & Population >= 30000, "medium",
                                   if_else(Population < 30000 & Population >= 1000, "small","rural")))) %>%
    relocate(Population, .before = Latitude) %>%
    relocate(PopSize, .before = Latitude) 

# Simplifying city names for map labels & dropdown menu
df <- df %>%
  unite("City", City:P, sep = ", ") %>%
  unite("NAPS", City:NAPS, sep = ", NAPS: ") 

write.csv(df, file = "www/ECCC2018_wideCombined.csv")


```

-----------------------------------------

## Comparison between tidy and wide

Currently, combining and saving data in the 'wide' format *significantly* reduces file size. For example, saving the combined 2018 NO2 and O3 hourly data in the wide format takes 11.8 mb, whereas saving the same data in the 'long' tidy format takes 126 mb. 

# Importing and Tidying Data

I wrote a quick function called `ECCCTidy` which tidies a single ECCC NAPS `.csv` file. In other words, it converts the ECCC 'matrix' layout into a 'long' layout where each row is the measurement of that specific pollutant at a given date-time and location. *Note* that all columns start with a capital letter to maintain consistency with the original NAPS dataset.

```{r, message = FALSE, warning = FALSE, error = FALSE}

ECCCTidy <- function(file, rows = Inf){
  
  # Getting pollutant from file name
  chem <- sub("\\_.*", "", file)
  
  
  # Skipping ECCC header when importing file
  df <- read_csv(file, skip = 7, n_max = rows)
  
  
  # Actually tidying ECCC file
  df <- df %>% rename_all(funs(gsub("\\..*","", make.names(names(df))))) %>%
    pivot_longer(
      cols = starts_with("H"),
      names_to = "Hour",
      values_to = chem)  %>%
    mutate(Date_time = anytime(paste(Date, str_sub(Hour, -2, -1), ":00"))) %>%
    select(-c(Date, Hour, Pollutant)) %>%
  relocate(Date_time, .before = chem) 
  
  # Combining City w/ NAPS to simplify Shiny selection
  
  df <- df %>%
    unite("City_NAPS", City:NAPS, sep = ", NAPS: ")
  
  df
}


```

## Transforming Data for Applet

I wrote another function which will combine tidied *O~3~* and *NO~2~* datasets,[^3] and calculate the *O~x~* value at each given time. All `-999` values are converted to `NA`, and therefore are not used in any subsequent plotting/calculations. I also added a `rows` input where you can specify the number of rows you want to import from the `.csv`. The default is `Inf` (read every row), but you can specify smaller numbers when testing stuff.

[^3]: Note this uses inner_join, so only O3 and NO2 values from stations found in BOTH datasets will be included.

```{r, message = FALSE, warning = FALSE, error = FALSE}
# Test data to combine 
NO2 <- ECCCTidy("NO2_2018.csv", rows = 2500)
O3 <- ECCCTidy("O3_2018.csv", rows = 2500)

# Row outline of function, will need to clean up and expand to include other pollutants (i.e. SO2)
ECCCCombine <- function(O3, NO2){
  df <- O3 %>% 
  inner_join(NO2) %>%
  na_if(-999) %>%
  mutate(Ox = O3 + NO2)
}


```

The `ECCCCombine` function is pretty basic right now, but in the future I hope to expand it so that :

-   Users can directly specify the ECCC files they want to use
-   Multiple pollutants can be included such as SO~2~, NO, and maybe stuff like PM~2.5~; essentially any ECCC formatted `.csv`.
-   Clean up the naming conventions to avoid duplication, etc.

The end result of this function will be saved as a `.csv` and *this will be the file uploaded to and used by the applet*.

Quickly plotting the combined ECCC data shows that the NAs are properly plotted as gaps in the timeseries, and that everything is kosher.

```{r, message = FALSE, echo = FALSE, warning = FALSE, fig.height = 4, fig.cap = "Example plots from datasets. (A) Plot at single NAPS station of O3, NO2, and Ox values. (B) O3 values for different stations. Note gaps in time series"}

library(ggpubr)

x <- ECCCCombine(O3, NO2)

#write.csv(x = x, file = "www/ECCCCombined_2018_test.csv")

p <- ggplot(subset(x, City_NAPS == c("St Johns, NAPS: 10102"))) +
      geom_line(aes(x = Date_time, y = O3, colour = "O3")) +
      geom_line(aes(x = Date_time, y = NO2, colour = "NO2")) +
      geom_line(aes(x = Date_time, y = Ox, colour = "Ox")) + 
      theme(legend.position="right") +
      scale_x_datetime()

q <- ggplot(subset(x, City_NAPS == c("St Johns, NAPS: 10102", "Mount Pearl, NAPS: 10401")), 
          aes(x = Date_time, y = O3)) +
      geom_line() +
      facet_grid(vars(City_NAPS)) +
      scale_x_datetime()


ggarrange(p,q,
          labels = c("A", "B"), 
          nrow = 2)

```



# Notes for the downstream stuff

-   As it stands, combining all of the ECCC data into columns is easier than a massively long .csv (plus if I want to be able to export datasets to excel, this is the way (will need to fix date though...)). However., the 'long' format is easier w/ ggplot2, so it may be a good idea to convert to 'long' the subset of data to be plotted to take advantage of ggplot2 features...

-   Will need to add a rolling average feature to applet
